---
title: "An Analysis Workflow with netmatchRI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Analysis Workflow with netmatchRI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`netmatchRI` combines matched design construction, diagnostics,
randomization-based inference, and sensitivity analysis for observational
network data.

## Data

The data frame should contain one row per unit, a binary treatment, baseline
covariates, and an outcome. The network may be supplied as an adjacency or
graph-distance matrix. This example uses a small deterministic draw from the
package generator.

```{r}
library(netmatchRI)

sim <- simulate_netmatch_example(seed = 20260821, n = 32)
dat <- sim$data
net_dist <- sim$net_dist

dim(dat)
table(dat$Z)
```

## Match

Construct a dual-penalty design. HiGHS is the default open-source solver;
Gurobi and GLPK can be selected with the `solver` argument.

```{r}
m <- netmatch(
  data = dat,
  treat = "Z",
  covariates = c("X1", "X2", "X3"),
  network = net_dist,
  method = "dual",
  kappa = 2
)

m
```

## Summary and diagnostics

The summary reports matched sample sizes, effective sample sizes, matched-set
sizes, balance, network distances, and solver status. `diagnose_match()` returns
the balance and network diagnostics as data frames.

```{r}
summary(m)

diag <- diagnose_match(m)
diag$covariate_balance
diag$network_summary
```

## Randomization-based inference

`RI_Naive()` uses matched-set variances only. `RI_Sensitivity()` allows
residual covariance between matched sets, while `RI_Design()` uses the
design-based bound.

```{r}
ri_naive <- RI_Naive(m, "Y")
ri_sens <- RI_Sensitivity(m, "Y", eta = 0.03, rho = 0.10)
ri_design <- RI_Design(m, "Y")

rbind(ri_naive$result, ri_sens$result, ri_design$result)
```

## Sensitivity

Evaluate a grid of dependence parameters and compute the critical curve for a
chosen test level.

```{r}
grid <- sensitivity_grid(
  m,
  "Y",
  eta = c(0, 0.03, 0.06),
  rho = seq(0, 1, by = 0.25)
)
grid$grid
plot_sensitivity(grid)

crit <- critical_sensitivity(m, "Y", rho = seq(0, 1, by = 0.25))
crit$curve
```

## Interpretation

The sensitivity grid shows how the p-value changes as `eta` increases the
magnitude of residual covariance and `rho` controls its decay with network
distance. The critical curve reports the smallest nonnegative `eta` at each
`rho` that reaches the selected test level. If the naive analysis is already
not significant, no positive critical boundary is reported.

```{r}
crit$interpretation
```
