---
title: "An Analysis Workflow with netmatchRI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Analysis Workflow with netmatchRI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`netmatchRI` provides network-constrained matching, diagnostics,
randomization-based inference, and sensitivity analysis for observational
network data. Its three matching methods are dual-penalty, single-penalty, and
covariate-only matching.

## Data

The data frame should contain one row per unit, a binary treatment, baseline
covariates, and an outcome. The network may be supplied as an adjacency or
graph-distance matrix. This example uses a reproducible simulated dataset with
a nonzero treatment effect.

```{r}
library(netmatchRI)

sim <- simulate_netmatch_example(seed = 90141, n = 60, beta_z = 2)
dat <- sim$data
net_dist <- sim$net_dist

dim(dat)
table(dat$Z)
```

## Match

Construct a dual-penalty design. HiGHS is the default open-source solver;
Gurobi and GLPK can be selected with the `solver` argument.

```{r}
m <- netmatch(
  data = dat,
  treat = "Z",
  covariates = c("X1", "X2", "X3"),
  network = net_dist,
  method = "dual",
  kappa = 2,
  estimand = "ATT"
)

m
```

## Summary and diagnostics

The summary reports matched sample sizes, effective sample sizes, matched-set
sizes, balance, network distances, and solver status. `diagnose_match()` checks
covariate balance and network distance within matched sets.

```{r}
summary(m)

diag <- diagnose_match(m)
diag$covariate_balance
diag$network_summary
```

## Compare matching designs

Pass any selected matching designs as a named list. The two plots reproduce the
paper's covariate balance and covariate similarity diagnostics. Set
`include_before = FALSE` to omit the original unmatched sample.

```{r}
m_covariate <- netmatch(
  dat, "Z", c("X1", "X2", "X3"), net_dist,
  method = "covariate"
)
m_single <- netmatch(
  dat, "Z", c("X1", "X2", "X3"), net_dist,
  method = "single", kappa = 2
)
designs <- list(
  "Dual-penalty" = m,
  "Covariate-only" = m_covariate,
  "Single-penalty" = m_single
)
plot_covariate_balance(designs, include_before = TRUE)
plot_covariate_similarity(designs, include_before = TRUE)
```

## Randomization-based inference

`RI_unadjusted()` ignores dependence across matched sets. `RI_adjusted()` supplies design-based covariance bounds between matched sets with user-specified sensitivity parameters eta and rho, while `RI_design()` uses the
design-based covariance bound. These procedures test the sharp null of no causal effect for every individual.

The `estimand = "ATT"`, `"ATC"`, and `"ATE"` options in `netmatch()` select
unit matching weights used by `matched_data()` and the balance diagnostics.
They do not change these sharp-null tests. The `weight_type` argument of the
RI functions selects the matched-set weights in the test statistic.
The defaults are `estimand = "ATT"` and `weight_type = "ns"`.

```{r}
ri_unadjusted <- RI_unadjusted(m, "Y")
ri_adjusted <- RI_adjusted(m, "Y", eta = 0.5, rho = 0.3)
ri_design <- RI_design(m, "Y")

rbind(ri_unadjusted$result, ri_adjusted$result, ri_design$result)
```

## Sensitivity

Evaluate a grid of dependence parameters and compute the critical curve for a
chosen test level.

```{r}
sens <- sensitivity_grid(
  m,
  "Y",
  eta = seq(0, 0.5, by = 0.1),
  rho = seq(0, 0.3, by = 0.05)
)
sens$grid
plot_sensitivity(sens)

crit <- critical_sensitivity(m, "Y", rho = seq(0, 1, by = 0.25))
crit
crit$summary
plot_sensitivity(crit)
```

## Interpretation

The sensitivity grid shows how the p-value changes as `eta` scales the
design-based covariance bounds and `rho` controls their decay with network distance.
The critical curve gives the value of `eta` at each `rho` for which the
p-value equals `alpha`. Within `[0, 1]`, values below
the curve retain rejection and values above it do not.

The summary reports the critical eta at `rho = 1` and the critical ratio.
If the baseline analysis is already nonsignificant, no critical boundary is
reported. See `?critical_sensitivity` for the formulas and boundary cases.
