---
title: "Using netmatchRI for Observational Network Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using netmatchRI for Observational Network Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`netmatchRI` implements dual-penalty matching and randomization-based inference
for observational network data with network dependence and interference.
## Inspect The Package

```{r, eval = FALSE}
library(netmatchRI)

ls("package:netmatchRI")
help(package = "netmatchRI")
?netmatch
```

## Generate Example Data

```{r, eval = FALSE}
sim <- simulate_netmatch_example()
dat <- sim$data
net_dist <- sim$net_dist

table(dat$Z)
summary(lm(Y ~ Z, data = dat))$coefficients
```

For your own observational network data, use a data frame with one row per
unit, a binary treatment column such as `Z`, covariates such as `X1`, `X2`,
`X3`, and an outcome column such as `Y` for randomization-based inference.

## Build The Dual-Penalty Design

```{r, eval = FALSE}
m_dual <- netmatch(
  data = dat,
  treat = "Z",
  covariates = c("X1", "X2", "X3"),
  network = net_dist,
  method = "dual",
  kappa = 2,
  solver = "highs"
)

m_dual
summary(m_dual)
```

The default `solver = "highs"` is the preferred open-source backend based on
package-local N = 300--500 benchmarks; this does not imply that HiGHS is
universally fastest. `solver = "auto"` tries Gurobi, HiGHS, then GLPK. The
defaults are a 90-second time limit and 0.01 relative MIP gap. At N = 500,
HiGHS may return a validated time-limited incumbent; `netmatch()` warns with
the actual status and available gap when this occurs.

## Diagnostics And Randomization-Based Inference

```{r, eval = FALSE}
diagnostics <- diagnose_match(m_dual)
diagnostics$covariate_balance
diagnostics$network_distance

ri_naive <- RI_naive(m_dual, "Y")
ri_decay <- RI_decay(m_dual, "Y", eta = 0.03, rho = 0.10)
ri_design <- RI_design(m_dual, "Y")

rbind(ri_naive$result, ri_decay$result, ri_design$result)
```

`RI_decay()` supports sensitivity analysis across network-dependence parameters,
while `RI_design()` provides the design-based analysis.

## Sensitivity Analysis

```{r, eval = FALSE}
sens <- netmatch_sensitivity(
  m_dual,
  "Y",
  eta = seq(0, 0.05, by = 0.01),
  rho = seq(0, 1, by = 0.05)
)

plot_sensitivity(sens, type = "pvalue")

crit <- critical_sensitivity(m_dual, "Y")
plot_sensitivity(crit, type = "critical")
```

## Comparison Designs

`method = "covariate"` and `method = "single"` are included as comparison
designs:

```{r, eval = FALSE}
m_cov <- netmatch(dat, "Z", c("X1", "X2", "X3"), net_dist,
                  method = "covariate", kappa = 2)

m_single <- netmatch(dat, "Z", c("X1", "X2", "X3"), net_dist,
                     method = "single", kappa = 2)
```
