---
title: "Using netmatchRI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using netmatchRI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`netmatchRI` supports network-constrained matching and randomization inference
for observational data with possible graph dependence. The main workflow uses
the dual-penalty matched design.

## Inspect The Package

```{r, eval = FALSE}
library(netmatchRI)

ls("package:netmatchRI")
help(package = "netmatchRI")
?netmatch
```

## Generate Example Data

```{r, eval = FALSE}
sim <- simulate_netmatch_example()
dat <- sim$data
net_dist <- sim$net_dist

table(dat$Z)
summary(lm(Y ~ Z, data = dat))$coefficients
```

For your own study, use a data frame with one row per unit, a binary treatment
column such as `Z`, covariates such as `X1`, `X2`, `X3`, and an outcome column
such as `Y` for inference.

## Build The Dual-Penalty Design

```{r, eval = FALSE}
m_dual <- netmatch(
  data = dat,
  treat = "Z",
  covariates = c("X1", "X2", "X3"),
  network = net_dist,
  method = "dual",
  kappa = 2,
  solver = "auto"
)

m_dual
summary(m_dual)
```

If Gurobi is installed, `solver = "gurobi"` will use it directly. Otherwise,
`solver = "auto"` falls back to the open-source GLPK backend.

## Diagnostics And Inference

```{r, eval = FALSE}
diagnostics <- diagnose_match(m_dual)
diagnostics$covariate_balance
diagnostics$network_distance

ri_naive <- RI_naive(m_dual, "Y")
ri_decay <- RI_decay(m_dual, "Y", eta = 0.03, rho = 0.10)
ri_design <- RI_design(m_dual, "Y")

rbind(ri_naive$result, ri_decay$result, ri_design$result)
```

## Sensitivity Analysis

```{r, eval = FALSE}
sens <- netmatch_sensitivity(
  m_dual,
  "Y",
  eta = seq(0, 0.05, by = 0.01),
  rho = seq(0, 1, by = 0.05)
)

plot_sensitivity(sens, type = "pvalue")

crit <- critical_sensitivity(m_dual, "Y")
plot_sensitivity(crit, type = "critical")
```

## Comparison Designs

`method = "covariate"` and `method = "single"` are included as comparison
designs:

```{r, eval = FALSE}
m_cov <- netmatch(dat, "Z", c("X1", "X2", "X3"), net_dist,
                  method = "covariate", kappa = 2)

m_single <- netmatch(dat, "Z", c("X1", "X2", "X3"), net_dist,
                     method = "single", kappa = 2)
```
